\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-07 paper.qmd
Updated PBMC descriptions

To determine whether the UMAP representation with the specific (hyper-)parameter choice suggested by @chen2023 preserves the data structure present in $9\text{-}D$ (@fig-model-pbmc-author-proj (b)), we visualize the model constructed with UMAP overlaid on the $9\text{-}D$ data. The @fig-NLDR-variety (a) shows three well-separated clusters with big separations. However, as shown in @fig-model-pbmc-author-proj, there is no big separation between three clusters in $9\text{-}D$. Therefore, the suggested UMAP representation (@fig-NLDR-variety (a)) does not accurately represent the structure of PBMC3k dataset in $9\text{-}D$.

@@

>>>>

To determine whether the UMAP representation with the specific (hyper-)parameter choice suggested by @chen2023 preserves the data structure present in $9\text{-}D$ (@fig-model-pbmc-author-proj (b)), we visualize the model constructed with UMAP overlaid on the $9\text{-}D$ data. The @fig-NLDR-variety (a) shows three well-separated clusters with big separations. However, as shown in @fig-model-pbmc-author-proj (a2), there is no big separation between three clusters in $9\text{-}D$. Therefore, the suggested UMAP representation (@fig-NLDR-variety (a)) does not accurately represent the structure of PBMC3k dataset.

@@

The MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated. It is found that tSNE (@fig-mnist-mse (a)) provide the most reasonable representation for the digit 1 dataset, showing universally best. However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap (@fig-tsne-best).

### 2024-09-07 paper.tex
Updated PBMC descriptions

To determine whether the UMAP representation with the specific
(hyper-)parameter choice suggested by \citet{chen2023} preserves the
data structure present in \(9\text{-}D\)
(Figure~\ref{fig-model-pbmc-author-proj} (b)), we visualize the model
constructed with UMAP overlaid on the \(9\text{-}D\) data. The
Figure~\ref{fig-NLDR-variety} (a) shows three well-separated clusters
with big separations. However, as shown in
Figure~\ref{fig-model-pbmc-author-proj}, there is no big separation
between three clusters in \(9\text{-}D\). Therefore, the suggested UMAP
representation (Figure~\ref{fig-NLDR-variety} (a)) does not accurately
represent the structure of PBMC3k dataset in \(9\text{-}D\).

@@

>>>>

To determine whether the UMAP representation with the specific
(hyper-)parameter choice suggested by \citet{chen2023} preserves the
data structure present in \(9\text{-}D\)
(Figure~\ref{fig-model-pbmc-author-proj} (b)), we visualize the model
constructed with UMAP overlaid on the \(9\text{-}D\) data. The
Figure~\ref{fig-NLDR-variety} (a) shows three well-separated clusters
with big separations. However, as shown in
Figure~\ref{fig-model-pbmc-author-proj} (a2), there is no big separation
between three clusters in \(9\text{-}D\). Therefore, the suggested UMAP
representation (Figure~\ref{fig-NLDR-variety} (a)) does not accurately
represent the structure of PBMC3k dataset.

@@

The MSE for different binwidths (\(a_1\)) using tSNE, UMAP, PHATE,
PaCMAP, and TriMAP with different (hyper-)parameter settings
(Figure~\ref{fig-pbmc-mse}) were calculated. It is found that tSNE
(Figure~\ref{fig-mnist-mse} (a)) provide the most reasonable
representation for the digit 1 dataset, showing universally best.
However, \gD{} representation shows a big non-linear cluster and a small
cluster with a small gap (Figure~\ref{fig-tsne-best}).
